---
title: "bro Connection Log Analysis"
author: "Michael Winn"
date: "October 21, 2016"
output: html_document
---

```{r parseInput, echo = FALSE, message = FALSE, warning = FALSE}

    knitr::opts_chunk$set(fig.width=12, 
                   fig.height=8,
                   echo=FALSE,
                   warning=FALSE,
                   options(scipen=999),
                   cache = TRUE)
    rm(list = ls())
    
    ### INSTALL AND LOAD REQUIRED PACKAGES #######################################################
    required.packages <- c("ggplot2", 
                           "psych", 
                           "dplyr", 
                           "igraph", 
                           "gdata", 
                           "reshape2", 
                           "scales",
                           "grid", 
                           "gridExtra",
                           "lattice"
                           # "devtools", 
                           # "multigraph",
                           # "rcharts"
                           )
    
    new.packages <- required.packages[!(required.packages %in% installed.packages()[,"Package"])]
    if(length(new.packages)) install.packages(new.packages)
    
    # Output stored in the 'files' variable to suppress output while minimizing objects;
    # the files object is reassigned in the next statement.
    files <- lapply(required.packages, require, character.only = TRUE, quietly = TRUE)
    
    
    ### READ FILE DATA ###########################################################################
    files <- list.files(path=".", pattern="\\.log$")
    data <- lapply(files, read.csv, sep="\t", header=TRUE, stringsAsFactor = FALSE)
    data <- do.call("rbind", data)

    ### CALCULATE DESCRIPTIVE VARIABLES ##########################################################
    tm_from <- format(as.POSIXlt(min(data$ts), origin = "1970-01-01"), "%d-%b-%Y %H:%M:%S")
    tm_to <- format(as.POSIXlt(max(data$ts), origin = "1970-01-01"), "%d-%b-%Y %H:%M:%S")
    tm_duration <- as.numeric(difftime(as.POSIXct(max(data$ts), origin = "1970-01-01"), 
                            as.POSIXct(min(data$ts), origin = "1970-01-01"), 
                            units = "hours"))

    ### SET DATA TYPES AND SYNTESIZE MISSING VECTORS #############################################
    
    # originator / responder IP address and port
    data$id.orig_h  <- as.factor(data$id.orig_h)
      # data$id.orig_p  <- as.factor(data$id.orig_p)
    data$id.resp_h  <- as.factor(data$id.resp_h)
      # data$id.resp_p  <- as.factor(data$id.resp_p)
    
    # protocol
    data$proto <- as.factor(data$proto)

    # service
    data$service[data $service == "-"] <- "UNK"
    data$service <- as.factor(data$service)
        
    # duration
    data$duration <- as.numeric(data$duration)
      data$duration[is.na(data$duration)] <- 0
      
    # total bytes
    data$orig_bytes <- as.numeric(data$orig_bytes)
      data$orig_bytes[is.na(data$orig_bytes)] <- 0
    
    data$resp_bytes <- as.numeric(data$resp_bytes)
      data$resp_bytes[is.na(data$resp_bytes)] <- 0
    
    data$total_bytes <- data$orig_bytes + data$resp_bytes
    
    # total ip_bytes
    data$orig_ip_bytes <- as.numeric(data$orig_ip_bytes)
      data$orig_ip_bytes[is.na(data$orig_ip_bytes)] <- 0
    
    data$resp_ip_bytes <- as.numeric(data$resp_ip_bytes)
      data$resp_ip_bytes[is.na(data$resp_ip_bytes)] <- 0
    
    data$total_ip_bytes <- data$orig_ip_bytes + data$resp_ip_bytes
      
    # total pkts
    data$orig_pkts <- as.numeric(data$orig_pkts)
      data$orig_pkts[is.na(data$orig_pkts)] <- 0
    
    data$resp_pkts <- as.numeric(data$resp_pkts)
      data$resp_pkts[is.na(data$resp_pkts)] <- 0
  
    data$total_pkts <- data$orig_pkts + data$resp_pkts
      
    # vlan
    data$vlan <- as.factor(data$vlan)
    
    drops <- c('orig_bytes', 'resp_bytes', 'orig_pkts', 'resp_pkts', 'orig_ip_bytes', 'resp_ip_bytes')
    data <- data[ , !(names(data) %in% drops)]

    
    ### FUNCTION DECLARATIONS ###########################################################################
    
    ## Function:    summarizeColumns
    #   Purpose:    calculate summary statistics on the columns of a data frame (data)
    #   Input:      Data frame, trimmed to the columns of interest (rows contain data records)
    #   Output:     Data frame -- columns of the input data frame are transposed to rows
    #               columns are n, mean, sd, median, trimmed, mad, min, max, range, skew, kurtosis, std err, 
    #               variance, std error of the mean, and conf interval
    
    summarizeColumns <- function (data, conf.int = .95, rm.nas = FALSE){
      
      if(rm.nas == TRUE) data <- data[complete.cases(data),]
      
      # data$resp_pkts[is.na(data$resp_pkts)] <- 0
      result <- describe(data)
      result$var <- apply(as.matrix(data), 2, FUN = var)
      result$ci <- result$se * qt((conf.int/2) + .5, result$n-1)
      
      return(result)
    }
    
    ## Function:    buildEdgeList
    #   Purpose:    Builds an edge list of unique vertices extracted from the connection log
    #   Input:      connLog as a data frame
    #   Output:     data frame  (id.orig_h, id.resp_h, proto, total_bytes, duration, vlan, service) 
    #               of connections between nodes, summarized by total number of bytes.
    buildEdgeList <- function(connlog) {
      ### Build edge list
      
      temp <- select(connlog, id.orig_h, id.resp_h, proto, total_ip_bytes, total_pkts, total_ip_bytes, duration, vlan, service)

      edges <- summarise(group_by(temp, id.orig_h, id.resp_h, vlan, proto, service), total_ip_bytes = sum(total_ip_bytes))
        edges$vlan <- as.factor(edges$vlan)
        edges$service <- as.factor(edges$service)
        edges <- subset(edges, !is.na(total_ip_bytes))
        
        # Set width to a scale of 1:10 (or the actual number if less than 10) based on total_ip_bytes
        # logsize <- length(connlog$total_ip_bytes)
        logsize <- length(levels(as.factor(connlog$total_ip_bytes)))
        
        if (logsize == 0) {
          edges$width = 0
        } else if (logsize > 10) {
          logsize <- 10
        }
        
        edges$width <- as.integer(with(edges,
                                       cut(total_ip_bytes,
                                           breaks = unique(quantile(total_ip_bytes, probs = 1:logsize/logsize))
                                           ),
                                  na.rm = TRUE, include.lowest = TRUE))
  
        edges$width[is.na(edges$width)] <- 1
        
        
        edges$vlan <- droplevels(edges$vlan)
        edges$service <- droplevels(edges$service)
        edges$proto <- droplevels(edges$proto)
        
        edges <- arrange(edges, proto, service)
         
      return(edges)
      
    }
    
    ## Function:    buildVertexList
    #   Purpose:    Builds a vertex list of unique vertices extracted from the edge list
    #   Input:      edgeList. A data frame consisting of at least an originator (id.orig_h) 
    #               and responder (id.resp_h) vector
    #   Output:     data frame  (IP, total_conn) of nodes and a count of their total number occcurrences.
    buildVertexList <- function(edgeList) {
      
      edgeList <- ungroup(edgeList)
      
      temp <- select(edgeList, id.orig_h, id.resp_h) 
    
      O <-  arrange(count(temp, id.orig_h), desc(n))
      O$id.orig_h <- as.character(O$id.orig_h)
    
      R <- arrange(count(temp, id.resp_h), desc(n))
      R$id.resp_h <- as.character(R$id.resp_h)
    
      names(O) <- c("src", "n")
      names(R) <- c("src", "n")
    
      nodes <- rbind(O, R)
      nodes <- aggregate(n ~ src, data = nodes, FUN = sum)
    
      names(nodes) <- c("IP", "total_conn")
    
      nodes$IP <- as.factor(nodes$IP)

      return(nodes)
    }
    
```

# Introduction and Research Questions

On a passively monitored network, using the connection logs from the bro network security montior.

1) Can the the communication flow be visualized in a way  that informs sensor placement and configuration?

2) Is is possible to identify anamolous communication between hosts (e.g., for further investigation)?

## Descriptive Summary of Data Set

Initial analysis of the bro connection logs contained in ``r files`` covered a period of time from ``r tm_from`` to ``r tm_to``
(approximately ``r round(tm_duration, 2)`` hours). 

Summary statistics of the overall data set, followed by lists of the IP Addresses with the Top 10 and Bottom 10 volume of connections are listed below:

```{r networkData, echo = FALSE, message = FALSE, warning = FALSE}
  ### Build edge and node lists
    edges <- buildEdgeList(data)
  
    nodes <- buildVertexList(edges)

  ### Calculate additional descriptive variables    
    num_nodes <- length(nodes$IP)
    num_conn <- length(edges$id.orig_h)

    # Embed the vertex.size attribute as a function of quantiles (width = 10) of total connections
    # nodes$vsize <- as.integer(with(nodes, 
    #                                cut(total_conn, 
    #                                    breaks = quantile(total_conn, probs = 0:10/10))), 
    #                           na.rm = TRUE, include.lowest = TRUE)
    # nodes$vsize[is.na(nodes$vsize)] <- 1
    # nodes <- mutate(nodes, vsize = ifelse(vsize < 1, 1, vsize))

    dataSummary <- summarizeColumns(select(data, 
                            duration,
                            total_bytes, 
                            total_ip_bytes, 
                            total_pkts)
                     )[c(2, 3, 9, 4, 8, 10)] # n, mean, var, sd, se, ci
    # [c(2, 3, 14, 4, 13,15)] # n, mean, var, sd, se, ci
    dataSummary
    
    head(arrange(nodes, desc(total_conn)), 10)
    head(arrange(nodes, total_conn), 10)
    
```

During the logging time period, ``r num_nodes`` and unique nodes with ``r num_conn`` unique connections, and a total of ``r as.numeric(sum(data$total_pkts))`` packets (``r humanReadable(sum(data$total_ip_bytes), standard = "SI")``) were caclulated from data recorded in the observed IP headers. 
The communications contained ``r length(unique(data$service))`` unique services among ``r length(unique(data$proto))`` protocols. 
Additionally, bro recorded `r length(unique(data$vlan))` unique VLAN tags.

bro measures originator and responder bytes from observed TCP sequence numbers.
Missed bytes are also derived from TCP sequence numbers and are used to estimate packet loss. 
A total of ``r humanReadable(sum(data$total_bytes), standard = "SI")`` were recorded from TCP headers, with ``r humanReadable(sum(data$missed_bytes), standard = "SI")`` missing.
The estimated packet loss is approximately ``r round(sum(data$missed_bytes)/sum(data$total_bytes)*100, 1)`%`.

## Exploratory data analysis

```{r summaryTimeline, echo = FALSE, message = FALSE, warning = FALSE}
  
  bin <- 3600
  d <- select(data, ts, total_pkts)
  d <- arrange(d, ts)
  d$dattm <- as.POSIXct(d$ts, origin = "1970-01-01")
  
  ggplot(d, aes(dattm)) +
    geom_histogram(binwidth = bin) +
    ggtitle("Connection Volume, Grouped by Hour") +
    labs(x = "Time", y = "Total Packets") +
    scale_x_datetime(labels = date_format("(%d) %H:%M"),
                     breaks = date_breaks("4 hours"),
                     minor_breaks = date_breaks("1 hour")) +
    theme(axis.text.x  = element_text(angle=45, hjust = 1, vjust = 1, size = 8))

```

The following plot illustrates the distribution of the dependent variables that will be analyzed in the following sections. In order to isoloate skew, the plot is faceted by protocol (``r levels(data$proto)``). Additionally, solid lines represent the population mean and dashed lines represent the population median. The distance between the vertical lines indicates the skew for each variable.

```{r summary, echo = FALSE, message = FALSE, warning = FALSE}
  g1 <- ggplot(data, aes(x=vlan)) +
    geom_bar(stat = "count") +
    labs(x = "VLAN ID" , y = "Frequency") +
    ggtitle("by VLAN") +
    theme(axis.text.x  = element_text(angle=45, hjust = 1, vjust = 1, size = 8))
  
  g2 <- ggplot(data, aes(x=service)) +
    geom_bar(stat = "count") +
    labs(x = "Service", y = "") +
    ggtitle("by Service") +
    theme(axis.text.x  = element_text(angle=45, hjust = 1, vjust = 1, size = 8))

  grid.arrange(g1, g2, ncol = 2, top = "Population Distribution of Independent Variables")
  
  g1 <- ggplot(data, aes(duration)) + 
    geom_histogram() + 
    geom_vline(xintercept=mean(data$duration), size = 1) + 
    geom_vline(xintercept=median(data$duration), size = 1, linetype = "longdash") +
    labs(x = "Duration (s)", y = "Frequency") +
    scale_x_log10(limits = c(-1, max(data$duration))) +
    facet_grid(proto ~ .) +
    theme(axis.text.x  = element_text(angle=45, hjust = 1, vjust = 1, size = 8))

  g2 <- ggplot(data, aes(total_pkts)) + 
    geom_histogram() +
    geom_vline(xintercept=mean(data$total_pkts), size = 1) +
    geom_vline(xintercept=median(data$total_pkts), size = 1, linetype = "longdash") +
    labs(x = "Total Packets", y = "") +
    scale_x_log10(limits = c(-1, max(data$total_pkts))) +
    facet_grid(proto ~ .) +
    theme(axis.text.x  = element_text(angle=45, hjust = 1, vjust = 1, size = 8))
  
  g3 <- ggplot(data, aes(total_ip_bytes)) + 
   geom_histogram() + 
   geom_vline(xintercept=mean(data$total_ip_bytes), size = 1) + 
   geom_vline(xintercept=median(data$total_ip_bytes), size = 1, linetype = "longdash") +
   labs(x = "Total Bytes (by IP)", y = "") +
   scale_x_log10(limits = c(-1, max(data$total_ip_bytes))) +
   facet_grid(proto ~ .) +
   theme(axis.text.x  = element_text(angle=45, hjust = 1, vjust = 1, size = 8))
 
  g4 <- ggplot(data, aes(total_bytes)) + 
   geom_histogram() + 
   geom_vline(xintercept=mean(data$total_bytes), size = 1) + 
   geom_vline(xintercept=median(data$total_bytes), size = 1, linetype = "longdash") +
   labs(x = "Total Bytes (by TCP Header)", y = "") +
   scale_x_log10(limits = c(-1, max(data$total_bytes))) +
   facet_grid(proto ~ .) +
   theme(axis.text.x  = element_text(angle=45, hjust = 1, vjust = 1, size = 8))
  
  grid.arrange(g1, g2, g3, ncol = 3, top = "Population Distribution of Dependent Variables")

```

The levelplot illustrates the interaction of services and vlans, and is non-directional. The intensity of the colors from grey to red represents the total packets.

```{r levelplot, fig.width=24, fig.height=16}
  d <- select(data, vlan, service, total_pkts)
  m <- acast(d, vlan ~ service, value.var = "total_pkts")

  levelplot(m, 
            col.regions = colorRampPalette(c("grey","yellow","orange","red"),space="rgb"),
            aspect = "iso",
            scales = list(x = list(rot = 45)),
            xlab = "vlan ID",
            ylab = "service", 
            main = "Total Packets Exchanged between VLANs and Services"
  )
  
```
   
# Analysis of Network Connections

*[TODO] Discussion of the overall network communication picture.* 

Vertices are sized according to the relative volume of connections (inbound and outbound). Edges are directional, and colored by protocol. Edges thicknesses reflect the relative number of bytes transferred.

```{r networkgraph, echo = FALSE, message = FALSE, warning = FALSE}
# 
#   ### Build the graph
#     colrs <- rainbow(length(levels(edges$proto)))
#     edges$colors <- colrs[edges$proto]
#     
#     g <- graph_from_data_frame(d = edges, vertices = nodes, directed = TRUE) # Builds the basic igraph object
#     
#     # various layout algorithms
#     
#     # l <- layout_randomly(g)
#     # l <- layout_nicely(g)
#     # l <- layout_on_grid(g)
#     # l <- layout_on_sphere(g)
#     l <- layout_with_fr(g)
#     # l <- layout_with_kk(g)
#     # l <- layout_with_mds(g)
#     # l <- layout_with_drl(g)
#     # l <- layout_with_lgl(g)
#     # l <- layout_with_dh(g)
#     # l <- layout_with_fr(g)
#     
#     l <- norm_coords(l, ymin = -10, ymax = 10, xmin = -10, xmax = 10)
#     
#     # par(mfrow = c(1, 1), mar = c(0, 0, 0, 0))
# 
#     plot(g,                                       # Configures the aesthetics of the plot
#        vertex.size = coreness(g),
#        vertex.label = V(g)$IP,                  # Vertex name should be the node name (IP address)
#        # vertex.label.family = "Times",
#        vertex.label.cex = .75,
#        vertex.label.degree = 1,
#        edge.width = E(g)$width*0.5,               # Edge weight is a function of the number of bytes transferred
#        edge.color = E(g)$colors,                  # Edge color is a protocol (ICMP, TCP, UDP)
#        edge.arrow.size =  E(g)$width*0.1,
#        edge.curved = TRUE, 
#        # layout = l,
#        main = "Network Connection Visualization"
#        )
#     
#     legend(x = -1,
#            y = -.9,
#            c("icmp", "tcp", "udp"),
#            pch = 21,
#            col = "#777777",
#            pt.bg = colrs,
#            pt.cex = 2,
#            cex = .8,
#            bty = "n",
#            ncol = 1,
#            title = "Protocol"
#            )
#     
#     
```

The following graph separates the network communications by protocol (e.g. `r levels(data$proto)`).

*[TODO] Discussion of the overall network communication picture.* 

```{r networkgraph_byproto, echo = FALSE, message = FALSE, warning = FALSE}
  
    for(prot in levels(data$proto)) {
      
      ### Build edge list
      e <- buildEdgeList(subset(data, proto == prot))
      n <- buildVertexList(e)
    
    colrs <- rainbow(length(levels(edges$service)))
    e$colors <- colrs[e$service]
  
      h <- graph_from_data_frame(d = e,
                                 vertices = n, # use the existing (unmodified) node list...
                                 directed = TRUE) 
      
      # h <- delete_vertices(h, degree(h) == 0)     # ...remove nodes that do not have corresponding edges
  
      l <- layout_with_fr(h)
    
      plot(h,                                       # Configures the aesthetics of the plot
         # vertex.size =  as.integer(V(g)$vsize),   # Vertex size is a function of the total number of connections
         vertex.size = coreness(h),
         # vertex.label = V(h)$name,                  # Vertex name should be the node name (IP address)
         vertex.label = NA,                  # Omit Vertex name
         # vertex.label.family = "Times",
         # vertex.label.cex = .75,
         # vertex.label.degree = 1,
         # vertex.shape = "pie",
         edge.width = E(h)$width*0.5,               # Edge weight is a function of the number of bytes transferred
         edge.color = E(h)$colors,                  # Edge color is by service (dns, http, ssh, etc.)
         # edge.arrow.size =  E(h)$width*0.1,
         edge.arrow.size =  0.1,
         edge.curved = TRUE,
         layout = l,
         
        
         frame = FALSE,
         main = paste("Network graph by Protocol (", prot, ")")
         ) 
      
      legend(x = -1,
       y = -1,
       unique(E(h)$service),
       pch = 21,
       col = "#777777",
       pt.bg = colrs,
       pt.cex = 2,
       cex = .8,
       bty = "n",
       ncol = 1,
       title = "Service"
      )
      
    }
```  

## Network Traffic Subgraph 

``` {r anovastats, echo = FALSE, message = FALSE, warning = FALSE}
   ### Set iterators and data subset
    index <- 0

    for (i in 1:20) {
      
      index <- index + 1
      index <- sample(1:length(nodes$IP), 1, replace = FALSE)  # uncomment for diagnostic mode: random selection
      thisnode <- as.character(nodes$IP[index])
      
      # thisnode <- "155.19.59.227"  # uncomment for diagnostic mode: single selection
      # thisnode <- "50.63.243.228"  # uncomment for diagnostic mode: single selection
      # thisnode <- "50.19.59.43"  # uncomment for diagnostic mode: single selection

      subdata <- subset(select(data, 
                               id.orig_h, 
                               id.resp_h, 
                               proto, 
                               service, 
                               duration, 
                               total_bytes, 
                               total_pkts, 
                               total_ip_bytes, 
                               vlan),
                    (id.orig_h == thisnode | id.resp_h == thisnode))
                    
      subdata$id.orig_h <- droplevels(subdata$id.orig_h)
      subdata$id.resp_h <- droplevels(subdata$id.resp_h)
      subdata$service <- droplevels(subdata$service)
      
      
      subdatastats <- summarizeColumns(subdata[,5:8])[c(2, 3, 14, 4, 13,15)]  # (n, mean, var, sd, se, ci) 
                                                                              # for (duration, total_bytes, total_pkts, total_ip_bytes)
      
      subdatastats
      
      ## Diagnostic plots
      g1 <- ggplot(subdata, aes(duration)) + 
        geom_histogram() +
        geom_vline(xintercept=mean(subdata$duration), size = 1) +   # subdata is bold: solid = mean / dashed = median
        geom_vline(xintercept=median(subdata$duration), size = 1, linetype = "longdash") +      
        geom_vline(xintercept=mean(data$duration), size = 0.5) +    # population is normal: solid = mean / dashed = median
        geom_vline(xintercept=median(data$duration), size = 0.5, linetype = "longdash") +
        scale_x_log10(limits = c(-1, max(subdata$duration))) +
        facet_grid(proto ~ .) +
        theme(axis.text.x  = element_text(angle=45, hjust = 1, vjust = 1, size = 8))
      
      g2 <- ggplot(subdata, aes(total_pkts)) + 
        geom_histogram() + 
        geom_vline(xintercept=mean(subdata$total_pkts), size = 1) +
        geom_vline(xintercept=median(subdata$total_pkts), size = 1, linetype = "longdash") +      
        geom_vline(xintercept=mean(data$total_pkts), size = 0.5) + 
        geom_vline(xintercept=median(data$total_pkts), size = 0.5, linetype = "longdash") +
        scale_x_log10(limits = c(-1, max(subdata$total_pkts))) +
        facet_grid(proto ~ .) +    
        theme(axis.text.x  = element_text(angle=45, hjust = 1, vjust = 1, size = 8))
  
      # g3 <- ggplot(subdata, aes(total_bytes)) + 
      #   geom_histogram() + 
      #   geom_vline(xintercept=mean(subdata$total_bytes), size = 1) +
      #   geom_vline(xintercept=median(subdata$total_bytes), size = 1, linetype = "longdash") +      
      #   geom_vline(xintercept=mean(data$total_bytes), size = 0.5) + 
      #   geom_vline(xintercept=median(data$total_bytes), size = 0.5, linetype = "longdash") +
      #   scale_x_log10(limits = c(-1, max(subdata$total_bytes))) +
      #   facet_grid(proto ~ .) +    
      #   theme(axis.text.x  = element_text(angle=45, hjust = 1, vjust = 1, size = 8))
      
      g4 <- ggplot(subdata, aes(total_ip_bytes)) + 
        geom_histogram() + 
        geom_vline(xintercept=mean(subdata$total_ip_bytes), size = 1) +
        geom_vline(xintercept=median(subdata$total_ip_bytes), size = 1, linetype = "longdash") +      
        geom_vline(xintercept=mean(data$total_ip_bytes), size = 0.5) + 
        geom_vline(xintercept=median(data$total_ip_bytes), size = 0.5, linetype = "longdash") +
        scale_x_log10(limits = c(-1, max(subdata$total_ip_bytes))) +
        facet_grid(proto ~ .) +
        theme(axis.text.x  = element_text(angle=45, hjust = 1, vjust = 1, size = 8))
      
        grid.arrange(g1, g2, g4, ncol = 3, top = paste("Index", index, ": Sample Distribution of Dependent Variables for ", thisnode))
  
      # [TODO] Add the overall median and 1st/3rd QTR lines
      g1 <- ggplot(subdata, aes(x = id.resp_h, y = duration)) +
        geom_boxplot() +
        labs(x = "", y = "Duration (s)") +
        geom_hline(yintercept=mean(subdata$duration), size = 1) +
        geom_hline(yintercept=median(subdata$duration), size = 1, linetype = "longdash") +      
        geom_hline(yintercept=mean(data$duration), size = 0.5) + 
        geom_hline(yintercept=median(data$duration), size = 0.5, linetype = "longdash") +
        facet_grid(proto ~ .) +    
        theme(axis.text.x  = element_text(angle=45, hjust = 1, vjust = 1, size = 8))
        
      
      g2 <- ggplot(subdata, aes(x = service, y = duration)) +
        geom_boxplot() + 
        labs(x = "", y = "") +
        geom_hline(yintercept=mean(subdata$duration), size = 1) +
        geom_hline(yintercept=median(subdata$duration), size = 1, linetype = "longdash") +      
        geom_hline(yintercept=mean(data$duration), size = 0.5) + 
        geom_hline(yintercept=median(data$duration), size = 0.5, linetype = "longdash") +
        facet_grid(proto ~ .) +    
        theme(axis.text.x  = element_text(angle=45, hjust = 1, vjust = 1, size = 8))
        
      
      # g3 <- ggplot(subdata, aes(x = id.resp_h, y = total_bytes)) +
      #   geom_boxplot() +
      #   labs(x = "Dest IP", y = "Total IP Bytes") +
      #   geom_hline(yintercept=mean(subdata$total_bytes), size = 1) +
      #   geom_hline(yintercept=median(subdata$total_bytes), size = 1, linetype = "longdash") +      
      #   geom_hline(yintercept=mean(data$total_bytes), size = 0.5) + 
      #   geom_hline(yintercept=median(data$total_bytes), size = 0.5, linetype = "longdash") +
      #   facet_grid(proto ~ .) +    
      #   theme(axis.text.x  = element_text(angle=45, hjust = 1, vjust = 1, size = 8))
        
      
      g4 <- ggplot(subdata, aes(x = service, y = total_ip_bytes)) +
        geom_boxplot() +
        labs(x = "Service", y = "") +
        geom_hline(yintercept=mean(subdata$total_ip_bytes), size = 1) +
        geom_hline(yintercept=median(subdata$total_ip_bytes), size = 1, linetype = "longdash") +      
        geom_hline(yintercept=mean(data$total_ip_bytes), size = 0.5) + 
        geom_hline(yintercept=median(data$total_ip_bytes), size = 0.5, linetype = "longdash") +
        facet_grid(proto ~ .) +    
        theme(axis.text.x  = element_text(angle=45, hjust = 1, vjust = 1, size = 8))
        
    
      grid.arrange(g1, g2, g4, ncol = 3, top = paste("Index", index, ": Analysis of node", thisnode))
      
  
  # [TODO] ANOVA
      # sub.result <- lm(subdata$duration ~ subdata$id.resp_h + subdata$service)  
      # summary(sub.result)
      # anova(sub.result)
      # 
      # 
      # sub.plot <- data.frame(Fitted = fitted(sub.result), Residuals = resid(sub.result), Treatment = subdata$service)
      # ggplot(sub.plot, aes(Fitted, Residuals, colour = Treatment)) + geom_point()
      
  ### Build subgraph edge and node lists
  
      subedges <- buildEdgeList(subdata)
      subnodes <- buildVertexList(subedges)
  
      colrs <- rainbow(length(levels(subedges$service)))
      subedges$colors <- colrs[subedges$service]
  
      s <- graph_from_data_frame(d = subedges, vertices = subnodes, directed = TRUE) # Builds the basic igraph object
  
      l <- layout_with_fr(s)
      l <- norm_coords(l, ymin = -10, ymax = 10, xmin = -10, xmax = 10)
      
      plot(s,                                       # Configures the aesthetics of the plot
         # vertex.size =  as.integer(V(g)$vsize),   # Vertex size is a function of the total number of connections
         vertex.size = coreness(s)*2,
         # vertex.label.family = "Times",
         vertex.label.cex = .75,
         vertex.label.degree = 1,
         # vertex.shape = "pie",
         # edge.color = E(s)$vlan,                  # Edge color is a VLAN
         edge.width = E(s)$width*0.5,               # Edge weight is a function of the number of bytes transferred
         edge.color = E(s)$colors,                  # Edge color is a protocol (ICMP, TCP, UDP)
         # edge.arrow.size =  E(s)$width*0.2,
         edge.arrow.size =  0.5,
         edge.curved = TRUE,
         layout = l,
         main = paste("Index", index, ": Subgraph of node", thisnode)
  
         )
      
      curve_multiple(s, start = 0.2)
  
      legend(x = -2,
             y = -.9,
             levels(subedges$service),
             pch = 21,
             col = "#777777",
             pt.bg = colrs,
             pt.cex = 2,
             cex = .8,
             bty = "n",
             ncol = 1,
             title = "Service"
             )
    }

```
# TO DO SECTION

```{r graphstats, echo = FALSE, message = FALSE, warning = FALSE}

# mean(nodes$total_bytes)
# sd(nodes$total_bytes)
# mean(edges$total_bytes)
# sd(edges$total_bytes)
# 
# reciprocity(g)
# diameter(g)
# 
# hist(degree(g))
# 


```

 
